---
title: "Assignment 1a (Python) - refactor obesity dataset code"
output: html_notebook
---

The below code downloads data from the obesity dataset from the [UC Irvine ML repository](https://doi.org/10.24432/C5H31Z). It then preprocesses the data, fits a machine-learning model and shows results. The coding style intentionally is not perfect.

In this assignment you'll refactor this code using the principles taught in the Reproducible Coding practices course. Things to consider are for example:

* Modularity
* Documentation
* Tests
* Packaging
* Dependencies
* Version control
* etc.

For the full assessment criteria see the Rubric. For full details on the assignment see the `README.md`

```{r installing and loading packages}
#install.packages('httr', repos='http://cran.us.r-project.org')
#install.packages('readr', repos='http://cran.us.r-project.org')
#install.packages('ggplot2', repos='http://cran.us.r-project.org')
#install.packages('randomForest', repos='http://cran.us.r-project.org')
#install.packages("dplyr", repos='http://cran.us.r-project.org')

library(httr)
library(readr)
library(ggplot2)
library(randomForest)
library(dplyr)
```

```{r loading the data}
## Define the URL for the data
url <- 'https://archive.ics.uci.edu/static/public/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition.zip'

## Download the data from the URL
httr::GET(url, write_disk('obesity.zip', overwrite = TRUE)) |> stop_for_status()

unzip('obesity.zip', exdir = './data')

## Load the data
obesity <- readr::read_csv('data/ObesityDataSet_raw_and_data_sinthetic.csv', show_col_types = FALSE)
```

```{r renaming the variables}
## Renaming according to the tidyverse styleguide
obesity <-  obesity %>%
  rename(
    gender = Gender,
    age = Age,
    height = Height,
    weight = Weight,
    family_history_obesity = family_history_with_overweight,
    freq_highcal = FAVC,
    freq_veg = FCVC,
    main_meals = NCP,
    cons_between = CAEC,
    smoking_status = SMOKE,
    cons_water = CH2O,
    cal_monitoring = SCC,
    freq_phys_activity = FAF,
    time_tech_use = TUE,
    cons_alcohol = CALC,
    mode_transport = MTRANS,
    obesity_class = NObeyesdad
  )

```

```{r create and categorize BMI}
## Create the BMI variable
obesity$bmi <- obesity$weight / (obesity$height)^2

plot(obesity$bmi, type = 'l')

## This can go, because you already have a variable with these categories
  ## Also the other variable is more specific
  ## It also is not used anywhere else

#obesity <- obesity %>%
  #mutate(
   # outcome = case_when(
      #bmi < 18.5 ~ "0_underweight",
      #bmi < 25   ~ "1_normal",
      #bmi < 30   ~ "2_overweight",
      #TRUE       ~ "3_obese"
   # )
 # )


```

```{r split the data}
## Split the data for the web part and the synthetic part
## Do this after creating the BMI variable
obesity_web <- obesity[1:498, ]
obesity_synthetic <- obesity[499:nrow(obesity), ]

```

```{r converting variables to numeric}
## Make the mode of transport variable binary
obesity_web <- obesity_web %>%
  mutate(
    mode_transport = if_else(
      mode_transport %in% c("Walking", "Bike"),
      1L,
      0L
    )
  )

obesity_web$mode_transport <- as.numeric(obesity_web$mode_transport)

## Make the gender variable numeric
obesity_web$gender <- ifelse(obesity_web$gender == "Male", 1L, 0L)

obesity_web$gender <- as.numeric(obesity_web$gender)

## Define the binary variables
binary_vars <- c(
  "family_history_obesity",
  "freq_highcal",
  "smoking_status",
  "cal_monitoring"
)

## Convert the binary variables
for (col in binary_vars) {
  obesity_web[[col]] <- ifelse(obesity_web[[col]] == "yes", 1L, 0L)
}

## Converting the ordinal variables to numeric
obesity_web$cons_between <- as.numeric(factor(obesity_web$cons_between, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) - 1

obesity_web$cons_alcohol <- as.numeric(factor(obesity_web$cons_alcohol, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) - 1
```

```{r}
## This can go, because these are just all variables.
  ## Also, they are still the 'wrong' names. 
# X_columns <- c('Age', 'Gender', 'family_history_with_overweight',
            #    'FAVC', 'FCVC', 'NCP', 'CAEC', 'SMOKE', 'CH2O', 'SCC', 'FAF', 'TUE',
            #    'CALC', 'MTRANS')
```

## Regression problem

```{r}
classes <- sort(unique(df_web$label))
freq <- table(df_web$label)
class_weight <- length(df_web$label) / (length(classes) * as.numeric(freq))
names(class_weight) <- names(freq)

print(classes)
print(class_weight)
```

```{r}
# Note: result can deviate from the python code due to random seed!
set.seed(42)
idx <- seq_len(nrow(df_web))

idx_train <- sample(idx, size = floor(0.7 * length(idx)))
idx_test <- setdiff(idx, idx_train)

X_train <- df_web[idx_train, X_columns]
X_test <- df_web[idx_test, X_columns]
y_train <- df_web$BMI[idx_train]
y_test <- df_web$BMI[idx_test]

sample_weights <- class_weight[df_web$label[idx_train]]
```

```{r}
reg <- lm(BMI ~ ., data = cbind(X_train, BMI = y_train))
reg <- randomForest(x = X_train, y = y_train, ntree = 100, nodesize = 5, maxnodes = 20)
```

```{r}
preds <- predict(reg, X_test)

plot(preds, y_test, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
plot(predict(reg, X_train), y_train, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
sample_weights_test <- class_weight[df_web$label[idx_test]]

mae_baseline <- sum(abs(y_test - rep(mean(y_train), length(y_test))) * sample_weights_test) / sum(sample_weights_test)
mae_model <- mean(abs(y_test - preds) * sample_weights_test)

cat(sprintf('Mean absolute error (baseline): %.3f\n', mae_baseline))
cat(sprintf('Mean absolute error: %.3f\n', mae_model))
```
